---
title: "Working Document - Final Project"
author: "Jiaqing Zhao"
date: "2/28/2018"
output:
  html_document: default
  pdf_document: default
---

#Part 1: Data

```{r}
# Prepare the data
library(ggplot2)
library(dplyr)
chile <- read.csv("electoral_chile.csv")
```


```{r}
# recoding head_hh
chile$head_hh <- ifelse(chile$head_hh == "Father", 1, ifelse(chile$head_hh =="Mother", 0 , NA))
# filter out unwanted observations
chile_young <- chile %>% filter(year_finish_hs == 2007) %>% na.omit(std_math) %>% na.omit(std_grades) %>% na.omit(std_verbal)
chile_young <- chile_young %>% filter(year > 1988 & year < 1991)

# recode parents education
chile_young$ed_father_score <- ifelse(chile_young$ed_father == "College" | chile_young$ed_father == "Incomeplete Collge", 2, ifelse(chile_young$ed_father == "No education", 0, 2))
chile_young$ed_mother_score <- ifelse(chile_young$ed_mother == "College" | chile_young$ed_mother == "Incomeplete Collge", 2, ifelse(chile_young$ed_mother == "No education", 0, 2))
chile_young$avg_par_edu <- (chile_young$ed_father_score + chile_young$ed_mother_score)/2

# change grades from continuous to categorical
chile_young$std_grades_decent <- ntile(chile_young$std_grades, 10) 
chile_young$std_verbal_decent <- ntile(chile_young$std_verbal, 10) 
chile_young$std_math_decent <- ntile(chile_young$std_math, 10) 
chile_young$std_history_decent <- ntile(chile_young$std_history, 10) 

# save new file
write.csv(chile_young,"newdataset_chile.csv")
```

Now we explore the NAs
```{r}

#filter for NAs
na_test <- chile %>% 
  filter(is.na(verbal))

#filter out NAs
test <- chile %>% 
  filter(!is.na(verbal))


hist(na_test$income_group,
     main = "Score NAs - Income Group Distribution",
     xlab = "Income Group")
hist(test$income_group,
     ylim=c(0,35000),
     main = "Score Excluded NAs - Income Group Distribution",
     xlab = "Income Group")
hist(chile$income_group,
     ylim=c(0,35000),
     main = "All data - Income Group Distribution",
     xlab = "Income Group")
```


#Part 2: Exploration of variables of interests

```{r}
# checking the differences between gender and grades
new <- chile_young %>% group_by(registered_vote) %>% summarise(avg_history_score = mean(std_history), avg_math_score = mean(std_math), avg_verbal_score = mean(std_verbal), avg_science_score = mean(std_science))

ggplot(new) + geom_bar(aes(x = registered_vote, y = avg_history_score), stat = "identity", position = "dodge", width = .4, fill = "#FF6666") + coord_flip() + theme(panel.background = element_rect(fill = "white", colour = "grey50"))# + geom_hline(yintercept = 54.5)

ggplot(new, aes(x = registered_vote, y = avg_math_score)) + geom_bar(stat = "identity", position = "dodge", width = .4, fill = "#FF6666") +  coord_flip() + theme(panel.background = element_rect(fill = "white", colour = "grey50"))

ggplot(new, aes(x = registered_vote, y = avg_verbal_score)) + geom_bar(stat = "identity", position = "dodge", width = .4, fill = "#FF6666") + coord_flip() + theme(panel.background = element_rect(fill = "white", colour = "grey50"))

ggplot(new, aes(x = registered_vote, y = avg_science_score)) + geom_bar(stat = "identity", position = "dodge", width = .4, fill = "#FF6666") + coord_flip() + theme(panel.background = element_rect(fill = "white", colour = "grey50"))
```

Students who register votes have higher grades. Females have higher grades in general.


```{r}
# Explore relationship between income group and ratio of registered votes within that group
new2 <- chile_young %>% group_by(income_group, registered_vote) %>% summarise(sum = n())
temp <- new2 %>% group_by(income_group) %>% summarise(sum_new = sum(sum)) #%>% summarise()
temp2 <- new2 %>% filter(registered_vote == 1)
temp3 <- merge(temp,temp2,by = "income_group")
temp3$ratio <- temp3$sum/temp3$sum_new
ggplot(temp3, aes(x = income_group, y = ratio)) + geom_bar(stat = "identity", position = "dodge", width = .4, fill= "#FF0000") + coord_flip() + theme(panel.background = element_rect(fill = "white", colour = "grey50"))# + geom_hline(yintercept = 54.5)
```

The higher the income group, the higher percent of student will register vote

#Part 3: Correlations and Models

```{r}
# variable selection based on corrplot
library(corrplot)
chile_3 <- chile_young %>% select(registered_vote, avg_par_edu, n_family, n_family_work, head_hh, income_group, std_grades, std_verbal, std_math, std_history,std_science, std_notas, female)
cor_chile <- cor(chile_3)
corrplot(cor_chile, method = "circle")
```
verbal correlates with grades, math but not history or science
history correlates with science but not other grades.

```{r}
# preliminary linear logit model
l <- glm(registered_vote ~ female * head_hh + std_history + std_verbal + income_group + avg_par_edu + as.factor(region) , family=binomial(link='logit'), data = chile_young)
summary(l)
```

```{r}
# update 1
l1 <- glm(registered_vote ~ as.factor(female) + std_history + std_verbal + income_group + as.factor(school_owners) + avg_par_edu  +  as.factor(region) ,family=binomial(link='logit'), data = chile_young)
summary(l1)
```


```{r}
# update 2
l2 <- lm(registered_vote ~ as.factor(female) + as.factor(std_history_decent) + as.factor(income_group) +  as.factor(std_verbal_decent) + as.factor(school_owners) + avg_par_edu  +  as.factor(region) , data = chile_young)
summary(l2)
```

Part 4: Outputs
```{r}
library(popbio)

# verbal score plot
logi.hist.plot(chile_young$std_verbal, chile_young$registered_vote, logi.mod = 1, boxp = FALSE, type = "hist", col ="gray")
# income effect
logi.hist.plot(chile_young$income_group, chile_young$registered_vote, logi.mod = 1, boxp = FALSE, type = "hist", col ="gray")
```

All significant. We can start with these variables: gender, head_hh, grades, and income_group.

```{r}
library(stargazer)
stargazer(l1)
```

# Appendix: Bayesian Hierarchical Models

```{r eval=F}
# takes too long to run
library(brms);options(mc.cores = parallel::detectCores())
fit1 <- brm(registered_vote ~ as.factor(female) + std_history + std_verbal + (1 + std_verbal|income_group) + (1+ as.factor(school_owners)) ,family=binomial(link='logit'), data = chile_young)
```


